---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ipaddress

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN status](https://www.r-pkg.org/badges/version/ipaddress)](https://CRAN.R-project.org/package=ipaddress)
[![R build status](https://github.com/davidchall/ipaddress/workflows/R-CMD-check/badge.svg)](https://github.com/davidchall/ipaddress/actions)
[![Coverage status](https://codecov.io/gh/davidchall/ipaddress/branch/master/graph/badge.svg)](https://codecov.io/gh/davidchall/ipaddress?branch=master)
<!-- badges: end -->

This package provides classes for working with IP addresses, inspired by the Python [ipaddress](https://docs.python.org/library/ipaddress.html) module.

Here are some of the features:

* Fully supports both **IPv4 and IPv6** address spaces
* **Efficiently stores** addresses in their native format (a sequence of bits)
* Calculations are performed in C++ for **improved performance**
* Classes are **compatible with the tidyverse**
* Functions for **generation and analysis of IP data**


## Installation

You can install the released version of ipaddress from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("ipaddress")
```

Or you can install the development version from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("davidchall/ipaddress")
```


## Usage

This package provides the `ip_address()` and `ip_network()` classes, which can be used as standalone vectors or as columns within data frames and tibbles.

```{r example, message=FALSE}
library(tidyverse)
library(ipaddress)

x <- tibble(
  address = ip_address(c("192.168.0.1", "2001:db8::8a2e:370:7334")),
  network = ip_network(c("192.168.100.0/22", "2001:db8::/48"))
)

x
```

Input character vectors are validated as they are parsed.
Invalid inputs raise a warning and are replaced with `NA`.

```{r input_validation}
ip_address(c("255.255.255.255", "255.255.255.256"))
```

Functions are provided to enable common tasks:

```{r}
mutate(x, ipv6 = is_ipv6(address), in_net = is_within(address, network))
```

---

Please note that the 'ipaddress' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
