---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ipaddress

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ipaddress)](https://CRAN.R-project.org/package=ipaddress)
[![R build status](https://github.com/davidchall/ipaddress/workflows/R-CMD-check/badge.svg)](https://github.com/davidchall/ipaddress/actions)
[![codecov.io](https://codecov.io/github/davidchall/ipaddress/coverage.svg?branch=master)](https://codecov.io/github/davidchall/ipaddress?branch=master)
<!-- badges: end -->

The ipaddress package provides classes for storing IP addresses and networks.
Its interface is inspired by Python's [`ipaddress`](https://docs.python.org/library/ipaddress.html) and [`cyberpandas`](https://cyberpandas.readthedocs.io) modules.

Here are some of the features:

* IP addresses are stored in their native representation (a sequence of bits) for **efficient data storage**
* Calculations are performed in C++ for **improved performance**
* Classes are **compatible with the tidyverse**
* Supports both **IPv4 and IPv6** address spaces


## Installation

You can install the released version of ipaddress from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("ipaddress")
```

Or you can install the development version from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("davidchall/ipaddress")
```


## Usage

This package provides the `ip_address` and `ip_network` classes, which can be used as standalone vectors or as columns within data frames and tibbles.

```{r example, message=FALSE}
library(tidyverse)
library(ipaddress)

tibble(
  address = ip_address(c("0.0.0.1", "192.168.0.1", "2001:db8::8a2e:370:7334")),
  network = ip_network(c("92.0.2.0/24", "192.168.100.0/22", "2001:db8::/48"))
)
```

Input character vectors are automatically validated as they are parsed.
Invalid inputs raise a warning and are replaced with `NA`.

```{r input_validation}
ip_address(c("0.0.0.0", "255.255.255.255", "255.255.255.256"))
```

